Creating Data for "Social inequality in mortality among adults and elderly in northern Sweden 1801--2013"\footnote{Version 2017-08-27, with a longer time period.}

### 2017-08-29 CreateData/creDat2.Rnw
from 1801

vb <- rbind(aduu, foli)
vb$period <- as.factor(vb$period)
x <- with(vb, table(period, hisclass, useNA = "ifany"))
y <- prop.table(x, margin = 1) * 100
print(xtable(round(y, 1), digits = 0, 
             caption = "Hisclass by time period, 1901--2013 (per cent)."))
@

@@

<<groupbirths, results = 'asis'>>=
vb <- vb[vb$birthdate > 1761 & vb$birthdate < 1974, ]
vb <- ageWin(vb, c(40, 90))
starts <- seq(1761, 1981, by = 10)
vb$cohort <- cut(vb$birthdate, starts, labels = starts[-length(starts)])
x <- with(vb, table(cohort, hisclass, useNA = "ifany"))
y <- prop.table(x, margin = 1) * 100
print(xtable(round(y, 1), digits = 0, caption = "Hisclass by birth cohort, 1801--1980 (per cent)."))
@

@@

vb <- vb[order(vb$id, vb$age), ]
source("../R/getCancer.R")
getCancer
vb <- getCancer(vb)
source("../R/getHeart.R")
getHeart
vb <- getHeart(vb)
vb$exposure <- with(vb, exit - enter)
vb0 <- vb
vb <- vb[!is.na(vb$hisclass), ] # same as 'na.omit(vb)' (I think)
first <- function(x) x[1]
hisclass40 <- with(vb, tapply(hisclass, id, first))
indx <- match(vb$id, names(hisclass40))
vb$hisclass40 <- hisclass40[indx]
##
vb$hisclass60 <- 0
hisclass60 <- with(vb[vb$age > 59, ], tapply(hisclass, id, first))
indx <- match(vb$id, names(hisclass60))
vb$hisclass60 <- hisclass60[indx]
komp <- is.na(vb$hisclass60)
vb$hisclass60[komp] <- vb$hisclass40[komp]
vb$hisclass40 <- factor(vb$hisclass40, labels = c("1+2", "3+4+5+6+8", "7+9+10+11+12"))
vb$hisclass60 <- factor(vb$hisclass60, labels = c("1+2", "3+4+5+6+8", "7+9+10+11+12"))
vb$hisclass <- factor(vb$hisclass, labels = c("1+2", "3+4+5+6+8", "7+9+10+11+12"))
## End cut ####################################################
levels(vb$period)[8] <- "1991-2001" ## Changed from "1991-2000", fob90 extended.
levels(vb0$period)[8] <- "1991-2001"
save(vb0, file = "../data/vb0.rda")
save(vb, file = "../data/vb.rda")
@

@@

vb0$hisclass[is.na(vb0$hisclass)] <- 0
vb0$poplink <- vb0$dataset == "poplink"
vb0Tab <- aggregate(vb0[, c("event", "exposure")], 
                   vb0[, c("sex", "age", "period", "cohort", 
                          "urban", "civst", "hisclass", "poplink")], 
                   FUN = sum)
## Is this necessary: (See above!)
##vbTab$hisclass40 <- factor(vbTab$hisclass40)
##vbTab$hisclass60 <- factor(vbTab$hisclass60)
vbTab$age <- factor(vbTab$age)
save(vbTab, file = "../data/vbTab.rda")
save(vb0Tab, file = "../data/vb0Tab.rda")
@

@@

<<lisaold>>=
lisaold <- age.window(lisaold, c(40, 90))
lisaold$exposure <- with(lisaold, exit - enter)
lisaold$educ <- as.factor(lisaold$SUN2000NIVA_OLD)
lisaold$income <- cut(lisaold$DISPINKPERSF, c(-600, 730, 958, 1270, 500000),
                      labels = c("lowest", "low", "high", "highest"))
loTab <- aggregate(lisaold[, c("event", "exposure")], 
                   lisaold[, c("sex", "age", "period", "urban", 
                               "civst", "income", "educ")],
                   FUN = sum)
save(loTab, file = "../data/loTab.rda")
save(lisaold, file = "../data/lisaold.rda")
@

@@

Now let us put on emph{partner} on {\bf lisaold}.

<<partner>>=
load("../../Data/partner.rda")
##females <- lisaold[lisaold$sex == "female", ]
##par.fem <- partner[partner$LINNEID1 %in% females$LINNEID | partner$LINNEID2 %in% females$LINNEID2, ]
partner$id1 <- pmin(partner$LINNEID1, partner$LINNEID2)
partner$id2 <- pmax(partner$LINNEID1, partner$LINNEID2)
partner <- unique(partner[, c("LINNEIDAR", "id1", "id2")])
@

>>>>

vb <- rbind(aduu, foli)
vb$period <- as.factor(vb$period)
x <- with(vb, table(period, hisclass, useNA = "ifany"))
y <- prop.table(x, margin = 1) * 100
print(xtable(round(y, 1), digits = 0, 
             caption = "Hisclass by time period, 1801--2013 (per cent)."))
@

@@

<<groupbirths, results = 'asis'>>=
vb <- vb[vb$birthdate > 1711 & vb$birthdate < 1974, ]
vb <- ageWin(vb, c(40, 90))
starts <- seq(1711, 1981, by = 10)
vb$cohort <- cut(vb$birthdate, starts, labels = starts[-length(starts)])
x <- with(vb, table(cohort, hisclass, useNA = "ifany"))
y <- prop.table(x, margin = 1) * 100
print(xtable(round(y, 1), digits = 0, caption = "Hisclass by birth cohort, 1801--1980 (per cent)."))
@

@@

vb <- vb[order(vb$id, vb$age), ]
source("../R/getCancer.R")
getCancer
vb <- getCancer(vb)
source("../R/getHeart.R")
getHeart
vb <- getHeart(vb)
vb$exposure <- with(vb, exit - enter)
vb0 <- vb
vb <- vb[!is.na(vb$hisclass), ] # same as 'na.omit(vb)' (I think)
first <- function(x) x[1]
hisclass40 <- with(vb, tapply(hisclass, id, first))
indx <- match(vb$id, names(hisclass40))
vb$hisclass40 <- hisclass40[indx]
##
vb$hisclass60 <- 0
hisclass60 <- with(vb[vb$age > 59, ], tapply(hisclass, id, first))
indx <- match(vb$id, names(hisclass60))
vb$hisclass60 <- hisclass60[indx]
komp <- is.na(vb$hisclass60)
vb$hisclass60[komp] <- vb$hisclass40[komp]
vb$hisclass40 <- factor(vb$hisclass40, labels = c("1+2", "3+4+5+6+8", "7+9+10+11+12"))
vb$hisclass60 <- factor(vb$hisclass60, labels = c("1+2", "3+4+5+6+8", "7+9+10+11+12"))
vb$hisclass <- factor(vb$hisclass, labels = c("1+2", "3+4+5+6+8", "7+9+10+11+12"))
## End cut ####################################################
levels(vb$period)[8] <- "1991-2001" ## Changed from "1991-2000", fob90 extended.
levels(vb0$period)[8] <- "1991-2001"
save(vb0, file = "../data2/vb0.rda")
save(vb, file = "../data2/vb.rda")
@

@@

vb0$hisclass[is.na(vb0$hisclass)] <- 0
vb0$poplink <- vb0$dataset == "poplink"
vb0Tab <- aggregate(vb0[, c("event", "exposure")], 
                   vb0[, c("sex", "age", "period", "cohort", 
                          "urban", "civst", "hisclass", "poplink")], 
                   FUN = sum)
## Is this necessary: (See above!)
##vbTab$hisclass40 <- factor(vbTab$hisclass40)
##vbTab$hisclass60 <- factor(vbTab$hisclass60)
vbTab$age <- factor(vbTab$age)
save(vbTab, file = "../data2/vbTab.rda")
save(vb0Tab, file = "../data2/vb0Tab.rda")
@

@@

<<lisaold>>=
lisaold <- age.window(lisaold, c(40, 90))
lisaold$exposure <- with(lisaold, exit - enter)
lisaold$educ <- as.factor(lisaold$SUN2000NIVA_OLD)
lisaold$income <- cut(lisaold$DISPINKPERSF, c(-600, 730, 958, 1270, 500000),
                      labels = c("lowest", "low", "high", "highest"))
loTab <- aggregate(lisaold[, c("event", "exposure")], 
                   lisaold[, c("sex", "age", "period", "urban", 
                               "civst", "income", "educ")],
                   FUN = sum)
save(loTab, file = "../data2/loTab.rda")
save(lisaold, file = "../data2/lisaold.rda")
@

@@

Now let us put on \emph{partner} on {\bf lisaold}. First we need to fix the file {\bf partner} so that we can distinguish 
females and males. We also remove homosexual couples: We want to check a female's status compared to her male partner's.

<<partner>>=
load("../../Data/partner.rda")
partner$id1 <- pmin(partner$LINNEID1, partner$LINNEID2)
partner$id2 <- pmax(partner$LINNEID1, partner$LINNEID2)
partner <- unique(partner[, c("LINNEIDAR", "id1", "id2")])
indx <- match(partner$id1, lisaold$LINNEID)
partner$sex1 <- lisaold$sex[indx]
indx <- match(partner$id2, lisaold$LINNEID)
partner$sex2 <- lisaold$sex[indx]
partner <- partner[!(is.na(partner$sex1) | is.na(partner$sex2)), ]
partner <- partner[partner$sex1 != partner$sex2, ]
dim(partner)
partner$female <- ifelse(partner$sex1 == "female", partner$id1, partner$id2)
partner$male <- ifelse(partner$sex1 == "male", partner$id1, partner$id2)
partner <- data.frame(AR = partner$LINNEIDAR,
                      female = partner$female,
                      male = partner$male)
@

Next, put on partners on the females of {\bf lisaold}:

<<putonpartner>>=
key <- paste(lisaold$LINNEID, lisaold$AR)
partner$key <- paste(partner$female, partner$AR)
indx <- match(key, partner$key)
lisaold$partner <- partner$male[indx]
@

Next, put on partner's \emph{education} and \emph{income}:

<<putinceduc>>=
males <- lisaold[lisaold$LINNEID %in% lisaold$partner, ]
females <- lisaold[!is.na(lisaold$partner), ]
indx <- match(females$partner, males$LINNEID)
females$m.educ <- males$educ[indx]
females$m.income <- males$income[indx]
@

Tabulating female's income and partner's income (\emph{m.income}):

<<tabminc>>=
x <- with(females, table(income, m.income))
x
@

Tabulating female's education and partner's education (\emph{m.educ}):

<<tabmeduc>>=
x <- with(females, table(educ, m.educ))
x
@

No comments!
